---
title: "1_RR_PrepareData"
author: "Nina-Alisa Kollakowski"
date: "`r Sys.Date()`"
output: html_document
---
## Demographic Data

```{r}
library(dplyr)
#load data from demographics questionnaire
load("Z:/Längsschnitt Selbst/Auswertung/T1/Fragebögen/Demographics Questionnaire/demographicData.RData")
demographicQuestionnaire <- overview %>%
  mutate(education_mother = Secondary_Education_Mother_num + Tertiary_Education_Mother_num,
         education_father = Secondary_Education_Father_num + Tertiary_Education_Father_num) %>%
  select(ID,
         education_mother,
         education_father,
         Age_Mother,
         Age_Father,
         Marital_Status_num,
         Siblings_num,
         Primary_Language_Child_num,
         Primary_Caretaker_num) %>%
  rename(id = ID)

#load age data of child at each time point
#T1: age of behavioral test, fNIRS, and sex
ages_T1 <- readxl::read_excel(path = "Z:/Längsschnitt Selbst/Daten/T1/VollständigkeitDatensätze_T1.xlsx", sheet = "Übersicht")
names(ages_T1) <- as.character(ages_T1[7,])
ages_T1 <- ages_T1[8:126,]
ages_T1 <- ages_T1 %>%
  select(`Alter T1_a (Tage)`, `Alter T1_b (Tage)`, Geschlecht, `VP-Nr.`) %>%
  rename(age_T1_spatialContingency = `Alter T1_a (Tage)`,
         age_T1_temporalContingency = `Alter T1_b (Tage)`,
         sex = Geschlecht,
         id = `VP-Nr.`)
ages_T1$id <- as.factor(ages_T1$id)
ages_T1$age_T1_spatialContingency <- as.numeric(ages_T1$age_T1_spatialContingency)
ages_T1$age_T1_temporalContingency <- as.numeric(ages_T1$age_T1_temporalContingency)
ages_T1$sex <- as.factor(ages_T1$sex)

#T2: age of behavioral test and EEG
ages_T2 <- readxl::read_excel(path = "Z:/Längsschnitt Selbst/Daten/T2/VollständigkeitDatensätze_T2.xlsx", sheet = "Übersicht")
names(ages_T2) <- as.character(ages_T2[4,])
ages_T2 <- ages_T2[5:111,]
ages_T2 <- ages_T2 %>%
  select(`Alter T2_a (Tage)`, `Alter T2_b (Tage)`, `VP Code`) %>%
  rename(age_T2_reaching = `Alter T2_a (Tage)`,
         age_sensoryAttenuation = `Alter T2_b (Tage)`,
         id = `VP Code`)
ages_T2$id <- as.factor(ages_T2$id)
ages_T2$age_T2_reaching <- as.numeric(ages_T2$age_T2_reaching)
ages_T2$age_sensoryAttenuation <- as.numeric(ages_T2$age_sensoryAttenuation)

#T4: age
ages_T4 <- readxl::read_excel(path = "Z:/Längsschnitt Selbst/Daten/T4/VollständigkeitDatensätze_T4.xlsx", sheet = "Übersicht")
names(ages_T4) <- as.character(ages_T4[4,])
ages_T4 <- ages_T4[5:104,]
ages_T4 <- ages_T4 %>%
  select(`Alter T4 (Tage)`, `VP Code`) %>%
  rename(age_T4 = `Alter T4 (Tage)`,
         id = `VP Code`)
ages_T4$id <- as.factor(ages_T4$id)
ages_T4$age_T4 <- as.numeric(ages_T4$age_T4)

demographicData <- merge(demographicQuestionnaire,
                         merge(ages_T1,
                               merge(ages_T2, 
                                     ages_T4,
                                     by = "id",
                                     all = TRUE),
                               by = "id",
                               all = TRUE),
                         by = "id",
                         all = TRUE)
rm(list = c("overview", 
            "ages_T1",
            "ages_T2",
            "ages_T4",
            "demographicQuestionnaire"))
```


## Spatial Contingency Data

```{r}
load("Z:/Längsschnitt Selbst/Auswertung/T1/Body Ownership_Behavioural/RScripts/allData.RData")
spatialContingencyData <- Result %>%
  group_by(id, view) %>%
  summarise(looking_s = sum(lookingDuration),
            legMovement_s = sum(legMovementDuration)) %>% #sum data by condition
  filter(view != "away") %>%
  tidyr::pivot_wider(names_from = view, values_from = c("looking_s", "legMovement_s")) %>%
  mutate(lookingScore = looking_s_mirrored/(looking_s_mirrored + looking_s_ego),
         legMovementScore = legMovement_s_mirrored/(legMovement_s_mirrored + legMovement_s_ego)) %>% #create contingency preference scores
  select(id, lookingScore, legMovementScore) %>%
  rename(spatialContingency_looking = lookingScore,
         spatialContingency_legs = legMovementScore)
rm(Result)
```
## Temporal Contingency Data - Looking

```{r}
library(dplyr)
load("Z:/Längsschnitt Selbst/Auswertung/T1/fNIRS/AttentionCoding/lookingData.RData")
temporalContingencyData_looking <- looking_data %>%
  select(ID, condition, mean_looking) %>%
  tidyr::pivot_wider(names_from = condition, values_from = mean_looking) %>% #create contingency preference score
  mutate(lookingScore = delayed/(delayed + online)) %>%
  select(ID, lookingScore) %>%
  rename(id = ID,
         temporalContingency_looking = lookingScore)
rm(looking_data)
```

## Temporal Contingency Data - fNIRS
```{r create time bins}
library(dplyr)
load("Z:/Längsschnitt Selbst/Auswertung/T1/fNIRS/RScripts/ROIData.RData")
n_bins <- 10
#create time bins and summarise data within these
fnirsData_binned <- fnirsData_ROI_long %>%
  mutate(bins = ntile(time, n = n_bins)) %>%
  group_by(subject, condition, hemisphere, bins) %>%
  summarise(HbO = mean(HbO),
            HbR = mean(HbR))
fnirsData_binned$bins <- as.factor(fnirsData_binned$bins)
fnirsData_binned$condition <- droplevels(fnirsData_binned$condition)
```

```{r analyze left HbO}
#ANOVA
timeAnova_left_HbO <- ez::ezANOVA(fnirsData_binned[fnirsData_binned$hemisphere == "l",],
                                  dv = HbO,
                                  wid = subject,
                                  within = c(bins, condition))
print(timeAnova_left_HbO)

#Post-Hoc T-Tests of time bins
bins_left_HbO <- data.frame(matrix(nrow = n_bins, ncol = 3))
colnames(bins_left_HbO) <- c("bins", "t", "p")
bins_left_HbO$bins <- 1:n_bins
for (i in levels(fnirsData_binned$bins)){
  currentTest <- fnirsData_binned %>%
    filter(bins == i,
           hemisphere == "l") %>%
    t.test(HbO ~ condition, data = ., paired = TRUE)
  bins_left_HbO[i, "p"] <- currentTest$p.value
  bins_left_HbO[i, "t"] <- currentTest$statistic
}

#compute difference score between conditions in significant time bins
left_HbO_diff <- fnirsData_binned %>%
  filter(hemisphere == "l",
         bins %in%  which(bins_left_HbO$p < 0.05)) %>%
  group_by(subject, condition) %>%
  summarise(m_HbO = mean(HbO)) %>%
  select(subject, condition, m_HbO) %>%
  tidyr::pivot_wider(names_from = condition, values_from = m_HbO) %>%
  mutate(leftHbOScore = delayed - online) %>%
  select(subject, leftHbOScore)

```

```{r analyze right HbO}
#ANOVA HbO right hemisphere
timeAnova_right_HbO <- ez::ezANOVA(fnirsData_binned[fnirsData_binned$hemisphere == "r",],
                                   dv = HbO,
                                   wid = subject,
                                   within = c(bins, condition))
print(timeAnova_right_HbO)

bins_right_HbO <- data.frame(matrix(nrow = n_bins, ncol = 3))
colnames(bins_right_HbO) <- c("bins","t", "p")
bins_right_HbO$bins <- 1:n_bins
for (i in levels(fnirsData_binned$bins)){
  currentTest <- fnirsData_binned %>%
    filter(bins == i,
           hemisphere == "r") %>%
    t.test(HbO ~ condition, data = ., paired = TRUE)
  bins_right_HbO[i, "p"] <- currentTest$p.value
  bins_right_HbO[i, "t"] <- currentTest$statistic
}

#compute difference score between conditions in significant time bins
right_HbO_diff <- fnirsData_binned %>%
  filter(hemisphere == "r",
         bins %in%  which(bins_left_HbO$p < 0.05)) %>%
  group_by(subject, condition) %>%
  summarise(m_HbO = mean(HbO)) %>%
  select(subject, condition, m_HbO) %>%
  tidyr::pivot_wider(names_from = condition, values_from = m_HbO) %>%
  mutate(rightHbOScore = delayed - online) %>%
  select(subject, rightHbOScore)

```

```{r analyze left HbR}
timeAnova_left_HbR <- ez::ezANOVA(fnirsData_binned[fnirsData_binned$hemisphere == "l",],
                                  dv = HbR,
                                  wid = subject,
                                  within = c(bins, condition))
print(timeAnova_left_HbR)
#no significant effects of condition
```

```{r analyze right HbR}

timeAnova_right_HbR <- ez::ezANOVA(fnirsData_binned[fnirsData_binned$hemisphere == "r",],
                                   dv = HbR,
                                   wid = subject,
                                   within = c(bins, condition))
print(timeAnova_right_HbR)
#no significant effects of condition
```
```{r clean-up}
temporalContingencyData <- merge(temporalContingencyData_looking, 
                                 merge(left_HbO_diff, right_HbO_diff, by = "subject", all = TRUE),
                                 by.x = "id", by.y = "subject", all = TRUE) %>%
  rename(temporalContingency_left = leftHbOScore,
         temporalContingency_right = rightHbOScore)
rm(list = c("n_bins", 
            "timeAnova_left_HbO", 
            "timeAnova_left_HbR", 
            "timeAnova_right_HbO",
            "timeAnova_right_HbR",
            "bins_left_HbO", 
            "bins_right_HbO",
            "temporalContingencyData_looking",
            "left_HbO_diff",
            "right_HbO_diff",
            "currentTest",
            "fnirsData_binned",
            "fnirsData_ROI_long",
            "i"))
```

## Reaching to the Self Data

```{r}
library(dplyr)
load("Z:/Längsschnitt Selbst/Auswertung/T2/Reaching/RScripts/allData.RData")
reachingToSelfData <- Result %>%
  filter(!is.na(Contact_Success)) %>%
  group_by(ID, Target_Location_general) %>%
  summarise(n_trials = n(),
            proportionContact = sum(Contact_Success)/n_trials) %>% #calculate proportion of trials with successful contact
  select(-n_trials) %>%
  tidyr::pivot_wider(names_from = Target_Location_general, values_from = proportionContact) %>%
  rename(id = ID,
         reaching_arm = arm,
         reaching_head = head)

rm(Result)
```

## Sensory Attenuation (EEG) Data
```{r}
sensoryAttenuationData_buffer <- read.csv("Z:/Längsschnitt Selbst/Auswertung/T2/EEG/Preprocessing_HAPPE/Trigger_SoundOnsetOhneAffe/ClusterBasedPermutationAnalysis_Fieldtrip/Significant_Clusters_ERP/meanClusterDifferences.csv")
sensoryAttenuationData <- sensoryAttenuationData_buffer %>%
  rename(id = ID,
         sensoryAttenuation_pos = posCluster,
         sensoryAttenuation_neg = negCluster)

rm(sensoryAttenuationData_buffer)
```

## Mirror Self-Recognition Data
```{r}
library(dplyr)
mirrorSelfRecognitionData_buffer <- readxl::read_excel("Z:/Längsschnitt Selbst/Auswertung/T4/MirrorSelfRecognition/Hauptkodierung_LinaMöllen/Coding_MSR_T4_Lina_final.xlsx") %>%
  select(ID, Recognition)
mirrorSelfRecognitionData <- mirrorSelfRecognitionData_buffer %>%
  rename(id = ID,
         mirrorSelfRecognition = Recognition)

rm(mirrorSelfRecognitionData_buffer)
```

## Video Self-Recognition Data
```{r}
library(dplyr)
library(readxl)

#load data from blind coding
videoSelfRecognitionData_blind <- read_excel("Z:/Längsschnitt Selbst/Auswertung/T4/VideoSelfRecognition/Coding_ExcelFiles/Hauptkodierung_AshleyChay/T4_SelfRecognition_Ashley_final.xlsx")

#load condition data
conditions <- read_excel("Z:/Längsschnitt Selbst/Auswertung/T4/VideoSelfRecognition/T4_SelfRecognition_Conditions.xlsx")

#merge together and only consider children that responded correctly in both conditions as recognizer
videoSelfRecognitionData <- merge(videoSelfRecognitionData_blind, conditions, by.x = c("ID", "Condition"), by.y = c("ID", "Video")) %>%
  select(ID, `self-recognition`, Condition.y) %>%
  tidyr::pivot_wider(names_from = Condition.y, values_from = `self-recognition`) %>%
  mutate(selfRecognition = ifelse(self == "y" & other == "n", 1, 0)) %>%
  select(ID, selfRecognition) %>%
  rename(id = ID,
         videoSelfRecognition = selfRecognition)

#clean-up
rm(conditions)
rm(videoSelfRecognitionData_blind)
```

## Body Size: Door Data
```{r}
library(dplyr)
load("Z:/Längsschnitt Selbst/Auswertung/T4/BodyAsObject/Door/RScripts/final_data_BodyAsObject_Door.RData")
bodySizeDoorData <- all_results %>%
  mutate(errors = ifelse(errors_binary == "Yes", 1, 0)) %>% #recode: 1 = error present, 0 = no errors
  select(ID, errors) %>%
  rename(id = ID,
         bodySizeDoor_errors = errors)

rm(all_results)
```

## Body Size: Toys Data
```{r}
library(dplyr)
load("Z:/Längsschnitt Selbst/Auswertung/T4/BodyAsObject/Toys/RScripts/allData.RData")
bodySizeToysData <- allData %>%
  mutate(error_binary = ifelse(errors == 0, 0, 1)) %>% #check if trial with error (1) or not (0)
  select(id, toy, error_binary) %>%
  group_by(id) %>%
  summarise(m_errors = mean(error_binary)) %>% #calculate proportion of trials with errors
  rename(bodySizeToys_errors = m_errors)

rm(allData)
```

## Body as Obstacle data
```{r}
library(dplyr)
load("Z:/Längsschnitt Selbst/Auswertung/T4/BodyAsObject/Obstacle/RScripts/allData.RData")
bodyAsObstacleData <- allData %>%
  mutate(error = ifelse(errors_binary == "Yes", 1, 0)) %>% #recode: 1 = error present, 0 = no errors
  group_by(ID) %>%
  summarise(m_errors = mean(error))%>% #calculate proportion of trials with errors
  rename(id = ID,
         bodyObstacle_errors = m_errors)

rm(allData)
```

## Self-Understanding Questionnaire data
```{r}
library(dplyr)
load("Z:/Längsschnitt Selbst/Auswertung/T4/Fragebögen/UCLA/UCLA.RData")
selfUnderstandingData <- cleandat %>%
  mutate(passed = (self_description_rel + self_recognition_rel)/2) %>%
  select(ID, passed) %>% #add proportion of passed items per scale and create average passed items
  rename(id = ID,
         selfUnderstanding_passed = passed)

rm(cleandat)
```

## Placing Sticker on Self Data
```{r}
library(dplyr)
placingStickerData_buffer <- readxl::read_excel("Z:/Längsschnitt Selbst/Auswertung/T4/Placing/Coding_Placing_T4_StefanieKrammer.xlsx")
placingStickerData <- placingStickerData_buffer %>%
  mutate(correctPlace = ifelse(Placement == "of", 1, ifelse(Placement == -99, NA, 0))) %>% #check if correct placement (of = own face) or missing data
  select(ID, Type, correctPlace) %>%
  na.omit() %>% #omit missing data
  group_by(ID) %>%
  summarise(m_correctTrials = mean(correctPlace)) %>% #create proportion of trials with correct placement
  rename(id = ID,
         placing_correct = m_correctTrials)

rm(placingStickerData_buffer)
```

## Join all data
```{r}
selfData <- merge(demographicData,
                  merge(bodyAsObstacleData,
                        merge(bodySizeDoorData,
                              merge(bodySizeToysData,
                                    merge(mirrorSelfRecognitionData,
                                          merge(placingStickerData,
                                                merge(reachingToSelfData,
                                                      merge(selfUnderstandingData,
                                                            merge(sensoryAttenuationData,
                                                                  merge(spatialContingencyData,
                                                                        merge(temporalContingencyData,
                                                                              videoSelfRecognitionData,
                                                                              by = "id",
                                                                              all = TRUE),
                                                                        by = "id",
                                                                        all = TRUE),
                                                                  by = "id",
                                                                  all = TRUE),
                                                            by = "id",
                                                            all = TRUE),
                                                      by = "id",
                                                      all = TRUE),
                                                by = "id",
                                                all = TRUE),
                                          by = "id",
                                          all = TRUE),
                                    by = "id",
                                    all = TRUE),
                              by = "id",
                              all = TRUE),
                        by = "id",
                        all = TRUE),
                  by = "id",
                  all = TRUE)

save(selfData, file = "joinedSelfData.RData")
```

