---
title: "2_AnalyzeData"
author: "Nina-Alisa Kollakowski"
date: "`r Sys.Date()`"
output: html_document
---

## Initializing
```{r}
library(dplyr)
library(mice)
library(psych)
#library(semTools)
library(lavaan)

nImputations <- 500
```

## Defining measures

```{r}
##defining measures----
implicitSelfMeasures <- c("spatialContingency_looking", 
                          "spatialContingency_legs", 
                          "temporalContingency_looking",
                          "temporalContingency_left",
                          "temporalContingency_right",
                          "reaching_arm",
                          "reaching_head",
                          "sensoryAttenuation_pos",
                          "sensoryAttenuation_neg")
explicitSelfMeasures <- c("mirrorSelfRecognition", 
                          "videoSelfRecognition", 
                          "bodySizeDoor_errors",
                          "bodySizeToys_errors", 
                          "bodyObstacle_errors", 
                          "selfUnderstanding_passed", 
                          "placing_correct")
continuousVariables <- c("spatialContingency_looking", 
                         "spatialContingency_legs", 
                         "temporalContingency_looking",
                         "temporalContingency_left",
                         "temporalContingency_right",
                         "reaching_arm",
                         "reaching_head",
                         "sensoryAttenuation_pos",
                         "sensoryAttenuation_neg",
                         "bodySizeToys_errors", 
                         "bodyObstacle_errors", 
                         "selfUnderstanding_passed", 
                         "placing_correct")
dichotomousVariables <- c("mirrorSelfRecognition", 
                          "videoSelfRecognition", 
                          "bodySizeDoor_errors")
```

##load dataset
```{r}
load("joinedSelfData.RData")
currentData <- selfData
rm(selfData)
nObservations = nrow(currentData)
```

##scale dataset
without scaling CFA was giving the warnings that "some observed variances are (at least) a factor 1000 times larger than others; use varTable(fit) to investigate"
```{r}
currentData_scaled <- currentData %>% 
  mutate_at(continuousVariables, 
            ~(scale(.) %>% as.vector)) %>% 
  select(-id)
```

##impute Data
```{r}
#impute 500 different datasets and set seed to make reproducible
imputedData <- mice(currentData_scaled, m = nImputations, print = FALSE, seed = 350)
save(imputedData, file = "imputedData.Rdata")
```

#load imputed dataset
```{r}
load("imputedData.Rdata")
```

## compute correlation matrix for implicit/explicit measures for each imputed dataset
use psych::mixedCor for this
```{r echo=FALSE}
#initialize lists of correlations and standard deviations
correlationData_implicit <- vector("list", nImputations)
correlationData_explicit <- vector("list", nImputations)
standardDeviations_implicit <- vector("list", nImputations)
standardDeviations_explicit <- vector("list", nImputations)

for (i in 1:nImputations){
  #extract current imputed dataset
  curDataset <- complete(imputedData, action = i)
  
  #compute standard deviations of measures as a matrix as needed for covariance matrix
  standardDeviations_implicit[[i]] <- diag(sapply(curDataset %>%
                                                    select(all_of(implicitSelfMeasures)),
                                                  sd)) #implicit
  
  standardDeviations_explicit[[i]] <- diag(sapply(curDataset %>%
                                                    select(all_of(explicitSelfMeasures)), 
                                                  sd)) #explicit
  
  #calculate correlations of implicit measures    
  correlationData_implicit[[i]] <- mixedCor(curDataset %>%
                                              select(all_of(implicitSelfMeasures)),
                                            c = intersect(implicitSelfMeasures, continuousVariables),
                                            d = intersect(implicitSelfMeasures, dichotomousVariables))$rho 
  
  #calculate correlations of explicit measures
  correlationData_explicit[[i]] <- mixedCor(curDataset %>%
                                              select(all_of(explicitSelfMeasures)),
                                            c = intersect(explicitSelfMeasures, continuousVariables),
                                            d = intersect(explicitSelfMeasures, dichotomousVariables))$rho 
}
```

## create covariance matrix from each correlation matrix
```{r}
#initialize lists of covariances
covarianceData_implicit <- vector("list", nImputations)
covarianceData_explicit <- vector("list", nImputations)

for (i in 1:nImputations){
  covarianceData_implicit[[i]] <- standardDeviations_implicit[[i]] %*% correlationData_implicit[[i]] %*% standardDeviations_implicit[[i]]
  
  covarianceData_explicit[[i]] <- standardDeviations_explicit[[i]] %*% correlationData_explicit[[i]] %*% standardDeviations_explicit[[i]]
}
```

## create average matrices for implicit/explicit measures
```{r}
covariance_implicit <- apply(simplify2array(covarianceData_implicit), c(1,2), mean)
rownames(covariance_implicit) <- implicitSelfMeasures
colnames(covariance_implicit) <- implicitSelfMeasures

covariance_explicit <- apply(simplify2array(covarianceData_explicit), c(1,2), mean)
rownames(covariance_explicit) <- explicitSelfMeasures
colnames(covariance_explicit) <- explicitSelfMeasures

correlation_implicit <- apply(simplify2array(correlationData_implicit), c(1,2), mean)
rownames(correlation_implicit) <- implicitSelfMeasures
colnames(correlation_implicit) <- implicitSelfMeasures

correlation_explicit <- apply(simplify2array(correlationData_explicit), c(1,2), mean)
rownames(correlation_explicit) <- explicitSelfMeasures
colnames(correlation_explicit) <- explicitSelfMeasures

```


## compute CFA of implicit/explicit measures
MLR estimator
evaluate fit using X2-test, RMSEA, SRMR, CFI
good fit if 3/4 indicators suggest good fit
X2 = non-significant
RMSEA < .08
SRMR <.11
CFI >.95
```{r CFA implicit}
implicitCFA.model <- 'implicit =~ spatialContingency_looking + 
                                  spatialContingency_legs + 
                                  temporalContingency_looking + 
                                  temporalContingency_left +
                                  temporalContingency_right + 
                                  reaching_arm +
                                  reaching_head +
                                  sensoryAttenuation_pos +
                                  sensoryAttenuation_neg'
implicitCFA <- cfa(implicitCFA.model, sample.cov = covariance_implicit, sample.nobs = nObservations) 
summary(implicitCFA)
fitMeasures(implicitCFA, fit.measures = c("rmsea", "srmr", "cfi"))
```
significant X2-test (p = 0.000) -> bad
RMSEA > .08 (0.151) -> bad
SRMR > .11 (0.123) -> bad
CFI < .95 (0.302) -> bad
all indicators show bad fit of CFA for implicit measures -> perform EFA
```{r}
explicitCFA.model <- 'explicit =~ mirrorSelfRecognition + 
                                  videoSelfRecognition + 
                                  bodySizeDoor_errors + 
                                  bodySizeToys_errors +
                                  bodyObstacle_errors + 
                                  selfUnderstanding_passed +
                                  placing_correct'
explicitCFA <- cfa(explicitCFA.model, sample.cov = covariance_explicit, sample.nobs = nObservations) 
summary(explicitCFA)
fitMeasures(explicitCFA, fit.measures = c("rmsea", "srmr", "cfi"))
```
significant X2-test (p = 0.019) -> bad
RMSEA > .08 (0.089) -> bad
SRMR < .11 (0.075) -> good
CFI < .95 (0.692) -> bad
3/4 indicators show bad fit of CFA for explicit measures -> perform EFA

##perform EFA for implicit measures
1) Bartlett's test to check if correlation matrix different from identity matrix (needs to be significant)
```{r Bartlett}
cortest.bartlett(correlation_implicit, n = nObservations) #significant -> proceed
```

2) KMO >.5, otherwise exclude variables with smallest KMO until criterion is reached
```{r KMO}
KMO(correlation_implicit) #KMO = 0.47, exclude temporalContingency_left (0.42)
correlation_implicit1 <- correlation_implicit[!rownames(correlation_implicit) == "temporalContingency_left",
                                              !colnames(correlation_implicit) == "temporalContingency_left"]
excludedFactors_implicit <- list("temporalContingency_left")

#run KMO again
KMO(correlation_implicit1) #KMO = 0.51 -> proceed

```

3) parallel analysis with reduced correlation matrix to decide number of factors

```{r}
fa.parallel(correlation_implicit1, n.obs = nObservations, n.iter = 1000, fm = "ml", fa = "fa")
```
parallel analysis suggests 3 factors      

4) run EFA with principal axis factoring with oblique (direct oblimin) rotation
```{r}
#run EFA with 3 factors as suggested by parallel analysis
efa.implicit <- fa(correlation_implicit1, nfactors = 3, n.obs = nObservations, 
                   rotate = 'oblimin', fm = 'pa')

#print results: cut-off of .4 as only loadings with >=.4 will be assigned to this factor
print.psych(efa.implicit, cut = .4, digits=3, sort=TRUE) 

#exclude variables that don't load on any factor with loading of >=.4
excludedFactors_implicit <- append(excludedFactors_implicit, c("temporalContingency_looking",
                                                               "sensoryAttenuation_neg",
                                                               "sensoryAttenuation_pos"))

```

5) assign variables to factors with loadings of >=.4
6) calculate factor scores


## perform EFA for explicit measures
1) Bartlett's test to check if correlation matrix different from identity matrix (needs to be significant)
```{r Bartlett}

```

2) KMO >.5, otherwise exclude variables with smallest KMO until criterion is reached
```{r KMO}
KMO(correlation_implicit1) #KMO = 0.51 -> proceed
```

3) parallel analysis with reduced correlation matrix to decide number of factors
fa.parallel(implicitSelfCorrelations, n.obs = nrow(currentData), n.iter = 1000, fm = "ml")

4) run EFA with principal axis factoring with oblique (direct oblimin) rotation
5) assign variables to factors with loadings of >=.4
6) calculate factor scores

## perform EFA for explicit measures
1) Bartlett's test to check if correlation matrix different from identity matrix (needs to be significant)
```{r Bartlett}
cortest.bartlett(correlation_explicit, n = nObservations) #significant -> proceed
```

2) KMO >.5, otherwise exclude variables with smallest KMO until criterion is reached
```{r KMO}
KMO(correlation_explicit) #KMO = 0.53 -> proceed
```
3) parallel analysis with reduced correlation matrix to decide number of factors
4) run EFA with principal axis factoring with oblique (direct oblimin) rotation
5) assign variables to factors with loadings of >=.4
6) calculate factor scores

## linear regressions of implicit -> explicit self




